---
title: "2. Outbreak-size distributions by quarantine strategy"
subtitle: "Measles in Grantz schools, real SMART per-capita contact matrices, R0 = 12"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    df-print: kable
    embed-resources: true
  gfm:
    echo: false        # GitHub version shows results only; code is in the .qmd
    toc: true
    toc-depth: 2
execute:
  warning: false
  message: false
---

How big do outbreaks get under each quarantine strategy? For every school,
vaccination level and one of 11 strategies (**high / medium / low-risk
quarantine days**), we look at the full distribution of final outbreak
sizes rather than a single probability.

- **Density curves** overlay the three schools, one panel per strategy.
- **Histograms** show the raw counts for one school at a time.
- **Tables** give the median, 95th percentile and P(size ≥ K).

::: {.callout-note}
Model code lives in `measles_model.R`. Simulations are cached in
`results/`, and PNG copies of every figure are saved to `figures/`.
:::

::: {.content-hidden when-format="markdown"}

## Setup

```{r setup}
source("measles_model.R")

K     <- 20                                   # threshold line and P(size >= K)
VACCS <- c(0.50, 0.60, 0.70, 0.80, 0.85)      # one set of figures each

STRATS <- c("21/21/21", "21/14/7", "21/7/7", "21/7/0",
            "14/14/0",  "14/7/7",  "14/7/0", "14/0/0",
            "7/7/0",    "7/0/0",   "0/0/0")

HIST_SCHOOLS <- SCHOOLS      # schools that get histogram figures
HIST_FACET_Y <- "free_y"     # "fixed" to compare bar heights across panels
```

## Simulations

Each school × vaccination × strategy combination is simulated once and reused
by every figure and table below.

```{r simulate}
design <- cross_vacc(strategy_table(STRATS), VACCS)
sims   <- run_or_load("distributions", simulate_design(design))
sims[, strategy := factor(strategy, levels = STRATS)]
```

```{r plot-functions}
## Density curves: all schools on one plot, one panel per strategy
plot_density <- function(v) {
  d <- sims[vacc == v]
  ggplot(d, aes(size, colour = school)) +
    geom_density(linewidth = 1.1, adjust = 1.1, key_glyph = "path") +
    geom_vline(xintercept = K, linetype = "longdash",
               linewidth = 0.6, colour = "grey45") +
    facet_wrap(~ strategy, ncol = 3, scales = "free_y") +
    scale_colour_manual(values = SCHOOL_COLORS, name = "School") +
    scale_x_continuous(limits = c(0, max(d$size)),     # same x-scale in every panel
                       expand = expansion(mult = c(0.01, 0.03))) +
    labs(x = "Final outbreak size", y = "Density",
         title = sprintf("Outbreak-size distribution by strategy, %s vaccinated",
                         vacc_lab(v)),
         subtitle = sprintf("Panels = high/medium/low quarantine days. Dashed line = %d cases", K)) +
    theme_measles(11) +
    theme(legend.position = "top", panel.grid.major.x = element_blank())
}

## Histogram for one school
plot_hist <- function(s, v) {
  ggplot(sims[school == s & vacc == v], aes(size)) +
    geom_histogram(binwidth = 2, boundary = 0, fill = SCHOOL_COLORS[[s]]) +
    facet_wrap(~ strategy, ncol = 3, scales = HIST_FACET_Y) +
    labs(x = "Outbreak size", y = "Number of runs",
         title = sprintf("%s school, %s vaccinated: outbreak-size distribution",
                         s, vacc_lab(v)),
         subtitle = "Panels = high/medium/low quarantine days") +
    theme_measles(11)
}

## Wide summary table: median, 95th percentile and P(>= K) per school
summary_table <- function(v) {
  s <- sims[vacc == v, .(Median = median(size),
                         P95    = as.numeric(quantile(size, 0.95)),
                         PK     = mean(size >= K)),
            by = .(school, strategy)]
  w <- dcast(s, strategy ~ school, value.var = c("Median", "P95", "PK"))
  setcolorder(w, c("strategy", unlist(lapply(SCHOOLS, function(x)
    paste0(c("Median_", "P95_", "PK_"), x)))))
  w
}

print_summary <- function(v) {
  w <- summary_table(v)
  print(knitr::kable(
    w, digits = c(0, rep(c(0, 0, 3), length(SCHOOLS))),
    col.names = c("Strategy (H/M/L)", unlist(lapply(SCHOOLS, function(x)
      paste(x, c("median", "95th pct", sprintf("P(\u2265%d)", K))))))))
}
```

:::

## Density curves, all schools

::: {.panel-tabset}

```{r density, results='asis', fig.width=11, fig.height=10}
for (v in VACCS) {
  cat("\n\n### ", vacc_lab(v), " vaccinated\n\n")
  print(save_fig(plot_density(v),
                 sprintf("fig_dist_all_schools_%02d.png", round(100 * v)), 11, 10))
  cat("\n\n")
  print_summary(v)
  cat("\n\n")
}
```

:::

## Histograms by school

```{r histograms, results='asis', fig.width=9, fig.height=10}
for (s in HIST_SCHOOLS) {
  cat("\n\n### ", s, "\n\n::: {.panel-tabset}\n\n")
  for (v in VACCS) {
    cat("\n\n#### ", vacc_lab(v), "\n\n")
    print(save_fig(plot_hist(s, v),
                   sprintf("fig_dist_%s_%02d.png", tolower(s), round(100 * v)), 9, 10,
                   dpi = 150))
    cat("\n\n")
  }
  cat("\n\n:::\n\n")
}
```
