---
title: "1. Tiered quarantine: outbreak probability tables"
subtitle: "Measles in Grantz schools, real SMART per-capita contact matrices, R0 = 12"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    df-print: kable
    embed-resources: true
  gfm:
    echo: false        # GitHub version shows results only; code is in the .qmd
    toc: true
    toc-depth: 2
execute:
  warning: false
  message: false
---

Tiered-quarantine experiment using the three real per-capita SMART contact
matrices (symmetrised for reciprocity, kept at real magnitude). R0 = 12 is
reached through a per-school transmission rate, not by scaling the matrix.

For each school we report **P(outbreak ≥ 10 / 20 / 50 cases)** across
vaccination coverage and quarantine strategies, written as
**high / medium / low-risk contact quarantine days**.

The diagnostic single-tier scenarios in the first table do not satisfy
high ≥ medium ≥ low; that constraint is applied to every other analysis.

**Significance.** Each scenario's full outbreak-size distribution is compared
with a reference scenario using a two-sided Wilcoxon rank-sum test: *No
quarantine* for the no-vaccination table, and *21/21/21* at the same coverage
for the tiered table. Stars: \* p < 0.05, \*\* p < 0.01, \*\*\* p < 0.001.
P-values are not adjusted for multiple comparisons.

::: {.callout-note}
Model code lives in `measles_model.R`. Simulations are cached in
`results/`; delete the CSV to re-run. For a quick test render with
`N_SIMS=200 quarto render 01_tiered_quarantine_tables.qmd`.
:::

::: {.content-hidden when-format="markdown"}

## Setup

```{r setup}
source("measles_model.R")

THRESHOLDS   <- c(10, 20, 50)                       # P(size >= K)
VACCS_TIERED <- c(0.50, 0.60, 0.70, 0.80, 0.90, 0.95)
REFERENCE    <- c(novac = "No quarantine", tiered = "21/21/21")  # for the Wilcoxon test
```

:::

## Transmission rate check

Each school gets its own transmission rate so that R0 = 12. The table checks
this against `epiworldR::compute_reproduction_number()`.

```{r r0-check}
r0_check <- rbindlist(lapply(SCHOOLS, function(s) data.table(
  School              = s,
  Students            = sum(ENROLLMENT[[s]]),
  `Transmission rate` = round(TRANS_RATE[[s]], 4),
  R0 = round(epiworldR::compute_reproduction_number(
         contact_matrix         = SCHOOL_CMAT[[s]],
         transmission_prob      = TRANS_RATE[[s]],
         infectious_period_days = INF_DAYS,
         check_reciprocity      = TRUE)$R, 2)
)))
knitr::kable(r0_check)
```

::: {.content-hidden when-format="markdown"}

## Scenarios

```{r design}
## A. No vaccination: which single tier is worth quarantining? (diagnostic)
design_novac <- strategy_table(c(
  "No quarantine"        = "0/0/0",
  "Only high-risk (21d)" = "21/0/0",
  "Only med-risk (21d)"  = "0/21/0",
  "Only low-risk (21d)"  = "0/0/21"))
design_novac[, vacc := 0]

## B. Tiered strategies (all high >= medium >= low) at each coverage
design_tiered <- cross_vacc(
  strategy_table(c("21/21/21", "21/14/14", "21/7/7", "21/0/0")),
  VACCS_TIERED)

design <- rbind(data.table(set = "novac",  design_novac),
                data.table(set = "tiered", design_tiered))
```

```{r simulate}
sims <- run_or_load("tables", simulate_design(design))
```

```{r tabulate}
## P(size >= K) for each threshold, one column per K
prob_table <- function(dt, by) {
  dt[, setNames(lapply(THRESHOLDS, function(k) mean(size >= k)),
                paste0("P(\u2265", THRESHOLDS, ")")),
     by = by]
}

tab <- prob_table(sims, by = c("set", "school", "strategy", "vacc"))

## Wilcoxon rank-sum test: each scenario's full outbreak-size distribution
## vs the reference scenario in the same table (same school and coverage)
wilcox_p <- function(x, ref) {
  p <- suppressWarnings(wilcox.test(x, ref, alternative = "two.sided",
                                    exact = FALSE)$p.value)
  if (is.finite(p)) p else 1      # identical distributions (all ties)
}

sig <- sims[, {
  ref_name <- REFERENCE[[set]]
  ref      <- size[strategy == ref_name]
  strats   <- unique(strategy)
  .(strategy = strats,
    p_value  = vapply(strats, function(st)
      if (st == ref_name) NA_real_ else wilcox_p(size[strategy == st], ref),
      numeric(1)))
}, by = .(set, school, vacc)]

format_p <- function(p) fifelse(is.na(p), "Reference",
                         fifelse(p < 0.001, "<0.001", sprintf("%.3f", p)))

## Stars are escaped so Markdown doesn't read them as bold/italic
sig_stars <- function(p) fcase(is.na(p), "Reference",
                               p < 0.001, "\\*\\*\\*",
                               p < 0.01,  "\\*\\*",
                               p < 0.05,  "\\*",
                               default = "")

tab <- sig[tab, on = .(set, school, vacc, strategy)]
tab[, `:=`(`Overall p-value` = format_p(p_value),
           Significance      = sig_stars(p_value))]
tab[, strategy := factor(strategy, levels = unique(design$strategy))]
setorder(tab, school, vacc, strategy)

PROB_COLS <- paste0("P(\u2265", THRESHOLDS, ")")
SIG_COLS  <- c("Overall p-value", "Significance")
```

:::

## Results by school

```{r results, results='asis'}
for (s in SCHOOLS) {
  cat("\n\n### ", s, "\n\n")

  cat("**No vaccination: which single tier to quarantine** (compared with no quarantine)\n\n")
  print(knitr::kable(
    tab[school == s & set == "novac", c("strategy", PROB_COLS, SIG_COLS), with = FALSE],
    col.names = c("Scenario", PROB_COLS, SIG_COLS),
    digits = 3, align = c("l", rep("r", length(PROB_COLS)), "r", "c")))

  cat("\n\n**Tiered strategies (H/M/L days) by vaccination coverage** (compared with 21/21/21)\n\n")
  t2 <- tab[school == s & set == "tiered"]
  t2[, Coverage := vacc_lab(vacc)]
  print(knitr::kable(
    t2[, c("Coverage", "strategy", PROB_COLS, SIG_COLS), with = FALSE],
    col.names = c("Coverage", "Strategy", PROB_COLS, SIG_COLS),
    digits = 3, align = c("l", "l", rep("r", length(PROB_COLS)), "r", "c")))
}
```
