---
title: "3. Which quarantine policy keeps outbreaks small?"
subtitle: "Measles in Grantz schools, real SMART per-capita contact matrices, R0 = 12"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    df-print: kable
    embed-resources: true
    fig-dpi: 150
  gfm:
    echo: false        # GitHub version shows results only; code is in the .qmd
    toc: true
    toc-depth: 2
execute:
  warning: false
  message: false
---

This is the main analysis. Every combination of school, vaccination coverage
and **high / medium / low-risk quarantine days** (with high e medium e low) is
simulated once, and all figures are computed from that single set of runs.

| Figure | Question |
|:--|:--|
| 1  | How risky is each school with minimal quarantine? |
| 2a | Where does each factor reduce outbreak risk the most? |
| 2b | Which combinations of factors make outbreaks likely? |
| 3  | Which quarantine combinations keep risk low? |
| 4  | Given high- and medium-risk days, how many low-risk days are worth it? |
| 5  | When outbreaks happen, how big do they get? |
| 6  | How does risk change with low-risk days, across thresholds? |

Plain-language conclusions, computed from the results, are at the end and are
also written to `results/conclusions.txt`.

::: {.callout-note}
Model code lives in `measles_model.R`. Simulations are cached in
`results/`, and PNG copies of every figure are saved to `figures/`.
:::

::: {.content-hidden when-format="markdown"}

## Setup

```{r setup}
source("measles_model.R")
library(rpart)            # decision tree for Figure 2 (ships with R)

## Design grid
VACCS <- c(0.60, 0.70, 0.80, 0.85)
HIGHS <- c(7L, 14L, 21L)
MEDS  <- c(0L, 7L, 14L, 21L)
LOWS  <- c(0L, 3L, 5L, 7L, 10L, 14L, 21L)

## Analysis choices
K_MAIN     <- 25L                  # an "outbreak" = MORE THAN K_MAIN cases
KS         <- c(10L, 35L, 50L)     # thresholds compared in Figure 6
TARGETS    <- c(0.10, 0.20, 0.30)  # acceptable outbreak risks (Figure 3)
OPT_SHARE  <- 0.80   # "optimal" = fewest days giving 80% of the possible benefit
MIN_GAIN   <- 0.02   # if quarantine can't cut risk by 2 points, optimal = 0 days
FIG3_LOWS  <- c(0L, 7L, 14L, 21L)  # low-risk days shown in Figure 3
TREE_DEPTH <- 3      # decision-tree depth (3 = at most 8 risk groups)
ADD_CAPTION <- FALSE # TRUE adds a methods caption under every saved figure
```

```{r plot-settings}
theme_set(theme_measles(12))

ylab_risk  <- sprintf("Chance of more than %d cases", K_MAIN)
VACC_LABS  <- vacc_lab(VACCS)
MIN_POLICY <- sprintf("%d / 0 / 0", min(HIGHS))

CAPTION <- sprintf(
  "Outbreak = more than %d cases. %s simulations per setting, each starting with one infected student.",
  K_MAIN, format(N_SIMS, big.mark = ","))

## save_fig() plus the optional caption
fig <- function(p, file, width, height) {
  if (ADD_CAPTION)
    p <- p + labs(caption = CAPTION) +
      theme(plot.caption = element_text(colour = "grey35", size = 9, hjust = 0))
  save_fig(p, file, width, height)
}
```

:::

## Simulations

```{r simulate}
design <- policy_grid(VACCS, HIGHS, MEDS, LOWS)
sims   <- run_or_load("policy_grid", simulate_design(design))

## Probability of an outbreak (> K_MAIN cases) at every design point
summ <- sims[, .(P = mean(size > K_MAIN)),
             by = .(school, vacc, q_high, q_med, q_low)]
summ[, vacc_lab := factor(vacc_lab(vacc), levels = VACC_LABS)]
```

`r nrow(design)` quarantine/vaccination settings per school ×
`r length(SCHOOLS)` schools × `r format(N_SIMS, big.mark = ",")` simulations
= `r format(nrow(sims), big.mark = ",")` runs.

::: {.callout-important}
## What "outbreak risk" means in this report

Each simulation starts with one infected student. An **outbreak** is a
simulation that ends with **more than `r K_MAIN` total cases**. The
**outbreak risk** of a setting is the share of its
`r format(N_SIMS, big.mark = ",")` simulations that ended this way, i.e.
P(final outbreak size > `r K_MAIN`). Figures 1 to 5 and the conclusions use
this definition; Figure 6 repeats it for several other thresholds.
:::

## 1. The problem

With minimal quarantine (only high-risk contacts quarantined), how likely is
an outbreak?

```{r fig1, fig.width=10, fig.height=4.5}
d1 <- summ[q_med == 0 & q_low == 0]

p1 <- ggplot(d1, aes(factor(q_high, levels = HIGHS), vacc_lab, fill = P)) +
  geom_tile(colour = "white", linewidth = 1) +
  geom_text(aes(label = pct(P, accuracy = 1), colour = P > 0.5),
            size = 4, show.legend = FALSE) +
  scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black")) +
  scale_fill_gradient(low = "#fff5eb", high = "#b30000", limits = c(0, 1),
                      labels = pct, name = ylab_risk) +
  facet_wrap(~ school) +
  labs(x = "High-risk quarantine (days)", y = "Students vaccinated",
       title = "1. Outbreak risk with only high-risk contacts quarantined",
       subtitle = "Darker = more dangerous") +
  theme(panel.grid = element_blank())

fig(p1, "fig1_baseline_risk.png", 10, 4.5)
```

## 2. What drives outbreak risk?

### 2a. Effect of each factor

For every setting, move **one** factor from its lowest to its highest allowed
value while everything else stays fixed. For each factor and school we show
the biggest reduction, and name the setting where it happens.

```{r fig2a-data}
FACTOR_TITLES <- c(vacc   = "Increasing vaccination",
                   q_high = "Increasing high-risk quarantine",
                   q_med  = "Increasing medium-risk quarantine",
                   q_low  = "Increasing low-risk quarantine")

## Risk at the lowest vs highest value of `var`, all else fixed
range_pairs <- function(var) {
  keys <- setdiff(c("school", "vacc", "q_high", "q_med", "q_low"), var)
  summ[, {
    o <- .SD[order(.SD[[var]])]
    if (.N < 2) NULL
    else .(from = o[[var]][1], to = o[[var]][.N], before = o$P[1], after = o$P[.N])
  }, by = keys][, factor := var]
}

## "80% vaccinated · quarantine high 21, medium 7 days"
context <- function(r, var) {
  q <- c(q_high = "high", q_med = "medium", q_low = "low")
  others <- setdiff(names(q), var)
  paste0(if (var != "vacc") sprintf("%s vaccinated \u00b7 ", pct(r$vacc)) else "",
         "quarantine ",
         paste(sprintf("%s %d", q[others], unlist(r[, ..others])), collapse = ", "),
         " days")
}

change <- function(r, var) {
  if (var == "vacc") sprintf("%s \u2192 %s", pct(r$from), pct(r$to))
  else sprintf("%d \u2192 %d days", r$from, r$to)
}

d2a <- rbindlist(lapply(names(FACTOR_TITLES), function(var) {
  pr <- range_pairs(var)
  pr[, effect := before - after]
  best <- pr[order(-effect)][, .SD[1], by = school][order(match(school, SCHOOLS))]
  rbindlist(lapply(seq_len(nrow(best)), function(k) {
    r <- best[k]
    data.table(factor = var, school = as.character(r$school),
               before = r$before, after = r$after,
               case = sprintf("%s (%s)\n%s", as.character(r$school),
                              change(r, var), context(r, var)))
  }))
}))
d2a[, factor_title := factor(FACTOR_TITLES[factor], levels = FACTOR_TITLES)]
d2a[, case := factor(case, levels = rev(unique(case)))]
d2a[, direction := ifelse(after <= before, "Risk goes down", "Risk goes up")]
```

```{r fig2a, fig.width=11, fig.height=10}
p2a <- ggplot(d2a, aes(y = case)) +
  geom_segment(aes(x = before, xend = after, yend = case, colour = direction),
               linewidth = 1.2,
               arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +
  geom_point(aes(x = before), colour = "grey45", size = 3) +
  ## labels sit outside the arrow ends so they never overlap
  geom_text(aes(x = before, label = pct(before, accuracy = 1),
                hjust = ifelse(after <= before, -0.4, 1.4)),
            colour = "grey30", size = 3.3) +
  geom_text(aes(x = after, label = pct(after, accuracy = 1), colour = direction,
                hjust = ifelse(after <= before, 1.3, -0.3)),
            size = 3.3, fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = c("Risk goes down" = "#2171b5",
                                 "Risk goes up"   = "#b30000"), name = NULL) +
  scale_x_continuous(labels = pct, limits = c(0, 1)) +
  facet_wrap(~ factor_title, ncol = 1, scales = "free_y") +
  labs(x = ylab_risk, y = NULL,
       title = "2a. Where does each factor reduce outbreak risk the most?",
       subtitle = "Biggest reduction found in each school. Grey dot = before, arrow = after. Each row names the setting where it happens") +
  theme(panel.grid.major.y = element_blank(),
        strip.text  = element_text(hjust = 0, size = 12),
        axis.text.y = element_text(size = 9, lineheight = 0.9))

fig(p2a, "fig2a_factor_effects.png", 11, 10)
```

### 2b. Risk groups

A decision tree looks at school, vaccination and all three quarantine lengths
**at the same time**, and repeatedly splits the settings into the groups that
differ most in outbreak risk.

```{r fig2b-data}
tree_dat <- summ[, .(P, school, vacc, q_high, q_med, q_low)]

fit <- rpart(P ~ school + vacc + q_high + q_med + q_low,
             data = tree_dat, method = "anova",
             control = rpart.control(maxdepth = TREE_DEPTH, cp = 0.005,
                                     minbucket = 6, xval = 0))

tree_dat[, leaf := rownames(fit$frame)[fit$where]]
leaf_ids <- unique(tree_dat$leaf)
paths    <- path.rpart(fit, nodes = as.numeric(leaf_ids), print.it = FALSE)

VAR_NAMES <- c("school", "vacc", "q_high", "q_med", "q_low")

## Describe the values of one variable inside a leaf, in words
describe <- function(v, x) {
  u <- sort(unique(x))
  if (v == "school") return(paste(as.character(u), collapse = " or "))
  rng <- if (length(u) == 1) as.character(u) else sprintf("%s\u2013%s", min(u), max(u))
  switch(v,
    vacc   = if (length(u) == 1) sprintf("%s vaccinated", pct(u))
             else sprintf("%s\u2013%s vaccinated", pct(min(u)), pct(max(u))),
    q_high = sprintf("high-risk quarantine %s days", rng),
    q_med  = sprintf("medium-risk quarantine %s days", rng),
    q_low  = sprintf("low-risk quarantine %s days", rng))
}

d2b <- rbindlist(lapply(leaf_ids, function(id) {
  conds <- paths[[id]][-1]
  vars  <- intersect(VAR_NAMES, unique(sub("\\s*(<|>|=).*$", "", conds)))
  rows  <- tree_dat[leaf == id]
  rule  <- if (length(vars) == 0) "All settings"
           else paste(vapply(vars, function(v) describe(v, rows[[v]]), ""), collapse = " + ")
  data.table(rule = rule, risk = mean(rows$P), n = nrow(rows))
}))

d2b[, rule_wrapped := vapply(strwrap(rule, 55, simplify = FALSE), paste, "", collapse = "\n")]
d2b[, rule_wrapped := make.unique(rule_wrapped)]
d2b[, rule_f := factor(rule_wrapped, levels = rule_wrapped[order(risk)])]
```

```{r fig2b, fig.width=11, fig.height=1.5 + 0.6 * nrow(d2b)}
p2b <- ggplot(d2b, aes(risk, rule_f, fill = risk)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = sprintf("%s  (%d settings)", pct(risk, accuracy = 1), n)),
            hjust = -0.05, size = 3.4) +
  scale_fill_gradient(low = "#fcbba1", high = "#99000d", limits = c(0, 1), guide = "none") +
  scale_x_continuous(labels = pct, breaks = seq(0, 1, 0.25), limits = c(0, 1.25)) +
  labs(x = sprintf("Average %s", tolower(ylab_risk)), y = NULL,
       title = "2b. Which combinations make outbreaks likely?",
       subtitle = "Groups found automatically by examining all factors together. Top = most dangerous") +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 10, lineheight = 0.9))

fig(p2b, "fig2b_risk_groups.png", 11, 1.5 + 0.6 * nrow(d2b))
```

## 3. Which quarantine combinations keep risk low?

Every quarantine combination, sorted into risk bands. Columns are medium /
low-risk days, grouped by high-risk days. Low-risk days are shown in weekly
steps to keep the figure readable.

```{r fig3, fig.width=11, fig.height=7}
## Shortest policy per target (used in the conclusions)
d3 <- rbindlist(lapply(TARGETS, function(tg) {
  summ[, {
    ok <- .SD[P <= tg][order(q_high + q_med + q_low, q_low, q_med)]
    if (nrow(ok) == 0) .(days = NA_integer_, label = "Not possible")
    else .(days  = ok$q_high[1] + ok$q_med[1] + ok$q_low[1],
           label = sprintf("%d / %d / %d", ok$q_high[1], ok$q_med[1], ok$q_low[1]))
  }, by = .(school, vacc_lab)][, target := tg]
}))

BAND_LABS <- c(sprintf("Under %s", pct(TARGETS[1])),
               sprintf("%s\u2013%s", pct(head(TARGETS, -1)), pct(TARGETS[-1])),
               sprintf("Over %s", pct(tail(TARGETS, 1))))
BAND_COLS <- colorRampPalette(c("#1a9850", "#fee08b", "#d73027"))(length(BAND_LABS))

d3p <- summ[q_low %in% FIG3_LOWS]
d3p[, band := cut(P, breaks = c(-Inf, TARGETS, Inf), labels = BAND_LABS)]
ML_LEVELS <- unique(d3p[order(q_med, q_low), sprintf("%d/%d", q_med, q_low)])
d3p[, ml := factor(sprintf("%d/%d", q_med, q_low), levels = ML_LEVELS)]
d3p[, high_lab := factor(sprintf("High-risk %d days", q_high),
                         levels = sprintf("High-risk %d days", HIGHS))]

p3 <- ggplot(d3p, aes(ml, vacc_lab, fill = band)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  scale_fill_manual(values = BAND_COLS, name = "Outbreak risk", drop = FALSE) +
  facet_grid(school ~ high_lab, scales = "free_x", space = "free_x") +
  labs(x = "Medium / low-risk quarantine days", y = "Students vaccinated",
       title = "3. Every quarantine combination: which ones keep risk low?",
       subtitle = "Each square is one policy. Green = safest. Read left to right to find the shortest policy that turns green") +
  theme(panel.grid = element_blank(), axis.text.x = element_text(size = 9))

fig(p3, "fig3_policy_options.png", 11, 7)
```

## 4. How many low-risk quarantine days are worth it?

For each fixed high/medium policy: the fewest low-risk days that give
`r pct(OPT_SHARE)` of the benefit low-risk quarantine can provide. Each cell
shows those days and the risk without  with them.

```{r fig4, fig.width=12, fig.height=5.8}
d4 <- summ[q_med > 0, {
  o     <- .SD[order(q_low)]
  risk0 <- o$P[o$q_low == 0]
  gain  <- cummax(risk0 - o$P)     # more days never counted as worse (smooths noise)
  G     <- max(gain)
  best  <- if (G < MIN_GAIN) 0L else o$q_low[which(gain >= OPT_SHARE * G)[1]]
  .(opt_low  = best,
    risk0    = risk0,
    risk_opt = risk0 - gain[o$q_low == best],
    risk_max = risk0 - G)
}, by = .(school, vacc, vacc_lab, q_high, q_med)]

hm_levels <- unique(d4[order(-q_high, -q_med), sprintf("High %d \u00b7 Med %d", q_high, q_med)])
d4[, hm := factor(sprintf("High %d \u00b7 Med %d", q_high, q_med), levels = rev(hm_levels))]

p4 <- ggplot(d4, aes(vacc_lab, hm, fill = opt_low)) +
  geom_tile(colour = "white", linewidth = 1) +
  geom_text(aes(label = ifelse(opt_low == 0, "Not\nworth it",
                               sprintf("%d days\n%s \u2192 %s", opt_low,
                                       pct(risk0, accuracy = 1), pct(risk_opt, accuracy = 1))),
                colour = opt_low > 10),
            size = 2.8, lineheight = 0.9, show.legend = FALSE) +
  scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black")) +
  scale_fill_gradient(low = "#f7fbff", high = "#08519c", limits = c(0, max(LOWS)),
                      name = "Optimal low-risk days") +
  facet_wrap(~ school) +
  labs(x = "Students vaccinated", y = "Fixed policy (days)",
       title = "4. How many low-risk quarantine days are worth it?",
       subtitle = sprintf(
         "Each cell: best number of low-risk days, and outbreak risk without \u2192 with them. \"Not worth it\" = risk drops less than %s",
         pct(MIN_GAIN))) +
  theme(panel.grid = element_blank())

fig(p4, "fig4_optimal_low.png", 12, 5.8)
```

## 5. How big do outbreaks get?

```{r fig5-helpers}
SIZE_LABS <- c("10 or fewer", "11\u201335", "36\u201350", "More than 50")
SIZE_COLS <- c("#fee5d9", "#fcae91", "#fb6a4a", "#a50f15")

## Share of runs in each size band, for the policy chosen per school/vacc
size_shares <- function(policy) {
  d <- sims[policy, on = .(school, vacc, q_high, q_med, q_low)]
  d[, size_cat := cut(size, breaks = c(-Inf, 10, 35, 50, Inf), labels = SIZE_LABS)]
  d <- d[, .N, by = .(school, vacc, size_cat)][, share := N / sum(N), by = .(school, vacc)]
  d[, vacc_lab := factor(vacc_lab(vacc), levels = VACC_LABS)]
  d[]
}

size_plot <- function(d) {
  ggplot(d, aes(vacc_lab, share, fill = size_cat)) +
    geom_col(width = 0.7, position = position_stack(reverse = TRUE)) +
    scale_fill_manual(values = SIZE_COLS, name = "Total cases", drop = FALSE) +
    facet_wrap(~ school) +
    labs(x = "Students vaccinated", y = "Share of simulations") +
    theme(panel.grid.major.x = element_blank())
}
```

::: {.panel-tabset}

### 5a. Minimal quarantine

```{r fig5a, fig.width=10, fig.height=4.8}
min_pol <- unique(summ[q_high == min(HIGHS) & q_med == 0 & q_low == 0,
                       .(school, vacc, q_high, q_med, q_low)])

p5a <- size_plot(size_shares(min_pol)) +
  scale_y_continuous(labels = pct) +
  labs(title = "5a. How big outbreaks get",
       subtitle = sprintf("Minimal quarantine (%s). Light = died out, dark = spread widely", MIN_POLICY))

fig(p5a, "fig5a_outbreak_size_minimal.png", 10, 4.8)
```

### 5b. Optimal quarantine

For each school and vaccination level, the optimal policy is the one with the
fewest total quarantine days that gets at least `r pct(OPT_SHARE)` of the
largest possible risk reduction (any high / medium / low combination). If
extra quarantine can't lower risk by `r pct(MIN_GAIN)`, the minimal policy is
kept.

```{r fig5b, fig.width=10, fig.height=5}
opt_pick <- function(dt) {
  dt    <- copy(dt)
  risk0 <- dt[q_high == min(HIGHS) & q_med == 0 & q_low == 0, P]
  best  <- min(dt$P)
  dt[, days := q_high + q_med + q_low]
  need <- if (risk0 - best < MIN_GAIN) 0 else OPT_SHARE * (risk0 - best)
  dt[risk0 - P >= need - 1e-12][order(days, P)][1, .(q_high, q_med, q_low)]
}

opt_pol <- summ[, opt_pick(.SD), by = .(school, vacc)]
opt_pol[, policy := sprintf("%d/%d/%d", q_high, q_med, q_low)]

d5b     <- size_shares(opt_pol[, !"policy"])
pol_lab <- opt_pol[, .(school, vacc_lab = factor(vacc_lab(vacc), levels = VACC_LABS), policy)]

p5b <- size_plot(d5b) +
  geom_text(data = pol_lab, aes(vacc_lab, 1, label = policy), inherit.aes = FALSE,
            vjust = -0.6, size = 3.3, fontface = "bold") +
  scale_y_continuous(labels = pct, limits = c(0, 1.08), breaks = seq(0, 1, 0.25)) +
  labs(title = "5b. How big outbreaks get with the optimal quarantine",
       subtitle = sprintf(paste0("Above each bar: optimal quarantine days (high/medium/low), the shortest policy\n",
                                 "giving %s of the largest possible risk reduction. Light = died out, dark = spread widely"),
                          pct(OPT_SHARE)))

fig(p5b, "fig5b_outbreak_size_optimal.png", 10, 5)
```

:::

## 6. Risk vs low-risk quarantine days, across thresholds

P(outbreak > K) against low-risk quarantine days, for all three schools on the
same plot. Colour family = school; darker shade = longer medium-risk
quarantine. Points are simulated values, lines are LOESS smooths, and ribbons
are 95% Wilson intervals.

```{r fig6-setup}
## Orange / green / blue families, lighter -> darker as medium days increase
CURVE_COLS <- c(
  Elementary_0 = "#F6C667", Elementary_7 = "#E69F00", Elementary_14 = "#B87900", Elementary_21 = "#7A4F00",
  Middle_0     = "#80CFA9", Middle_7     = "#009E73", Middle_14     = "#00734F", Middle_21     = "#004D35",
  High_0       = "#80B8E6", High_7       = "#0072B2", High_14       = "#005080", High_21       = "#003450")

threshold_plot <- function(K) {
  d <- sims[, {
    x  <- sum(size > K); n <- .N; ci <- wilson(x, n)
    .(P = x / n, lo = ci$lo, hi = ci$hi)
  }, by = .(school, vacc, q_high, q_med, q_low)]
  d[, curve := paste(school, q_med, sep = "_")]

  dir.create(RESULTS_DIR, showWarnings = FALSE)
  fwrite(d, file.path(RESULTS_DIR, sprintf("table_grid_p%d_all_schools.csv", K)))

  ggplot(d, aes(q_low, P, colour = curve, group = curve)) +
    geom_vline(xintercept = c(7, 14, 21), colour = "grey88", linewidth = 0.4) +
    geom_ribbon(aes(ymin = lo, ymax = hi, fill = curve),
                alpha = 0.06, colour = NA, show.legend = FALSE) +
    geom_smooth(method = "loess", formula = y ~ x, se = FALSE,
                span = 0.9, linewidth = 1.1) +
    geom_point(size = 1.4, alpha = 0.8) +
    facet_grid(vacc ~ q_high, labeller = labeller(
      vacc   = function(v) paste0(as.numeric(v) * 100, "% vaccinated"),
      q_high = function(h) paste0("High = ", h, " days"))) +
    scale_colour_manual(values = CURVE_COLS, breaks = names(CURVE_COLS),
                        labels = sub("_", ", medium ", names(CURVE_COLS)),
                        name = NULL, guide = guide_legend(nrow = 3, byrow = TRUE)) +
    scale_fill_manual(values = CURVE_COLS, guide = "none") +
    scale_x_continuous(limits = c(0, 21), breaks = c(0, 7, 14, 21),
                       expand = expansion(mult = c(0.01, 0.02))) +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
    labs(x = "Low-risk quarantine days", y = sprintf("P(outbreak > %d)", K),
         title = sprintf("6. Probability of outbreak > %d across school types", K),
         subtitle = "Darker shades = longer medium-risk quarantine") +
    theme_measles(11) +
    theme(legend.position = "top", legend.box = "vertical",
          panel.grid.major.x = element_blank())
}
```

::: {.panel-tabset}

```{r fig6, results='asis', fig.width=12, fig.height=9}
for (K in KS) {
  cat("\n\n### K = ", K, "\n\n")
  print(fig(threshold_plot(K), sprintf("fig6_grid_p%d_all_schools.png", K), 12, 9))
  cat("\n\n")
}
```

:::

## Supplement: every quarantine combination, per school

::: {.panel-tabset}

```{r supplement, results='asis', fig.width=11, fig.height=10}
summ[, high_lab := factor(paste0("High-risk = ", q_high, " days"),
                          levels = paste0("High-risk = ", HIGHS, " days"))]

for (s in SCHOOLS) {
  cat("\n\n### ", s, "\n\n")
  ps <- ggplot(summ[school == s],
               aes(factor(q_low, levels = LOWS), factor(q_med, levels = MEDS), fill = P)) +
    geom_tile(colour = "white", linewidth = 0.6) +
    geom_text(aes(label = pct(P, accuracy = 1), colour = P > 0.5),
              size = 2.6, show.legend = FALSE) +
    scale_colour_manual(values = c(`TRUE` = "white", `FALSE` = "black")) +
    scale_fill_gradient(low = "#fff5eb", high = "#b30000", limits = c(0, 1),
                        labels = pct, name = ylab_risk) +
    facet_grid(paste(vacc_lab, "vaccinated") ~ high_lab) +
    labs(x = "Low-risk quarantine (days)", y = "Medium-risk quarantine (days)",
         title = sprintf("Supplement: %s school, every quarantine combination", s),
         subtitle = "Empty cells are not allowed (high \u2265 medium \u2265 low)") +
    theme(panel.grid = element_blank())
  print(fig(ps, sprintf("supp_grid_%s.png", tolower(s)), 11, 10))
  cat("\n\n")
}
```

:::

## Conclusions

```{r conclusions}
out   <- character()
add   <- function(...) out <<- c(out, sprintf(...))
ascii <- function(x) gsub("\u00b7", "-", gsub("\u2192", "->", x))   # plain-text file

add("CONCLUSIONS  (outbreak = more than %d cases)", K_MAIN)
add("")

## 1. Riskiest school
base <- summ[q_high == min(HIGHS) & q_med == 0 & q_low == 0,
             .(P = mean(P)), by = school][order(-P)]
add("1. THE PROBLEM")
add("   With minimal quarantine, %s school is the riskiest (average %s)",
    as.character(base$school[1]), pct(base$P[1], 1))
add("   and %s school the least risky (average %s).",
    as.character(last(base$school)), pct(last(base$P), 1))
add("")

## 2. Levers
add("2. WHAT MATTERS MOST")
for (v in names(FACTOR_TITLES)) {
  add("   %s, biggest reduction:", FACTOR_TITLES[[v]])
  for (k in seq_len(nrow(d2a[factor == v]))) {
    r <- d2a[factor == v][k]
    add("     %s: risk %s -> %s  (%s)", r$school, pct(r$before, 1), pct(r$after, 1),
        ascii(sub("^[^(]*\\(", "", sub("\\)\n", ": ", as.character(r$case)))))
  }
}
g <- d2b[order(-risk)]
add("   Most dangerous combination: %s -> average risk %s.", g$rule[1], pct(g$risk[1], 1))
add("   Safest combination: %s -> average risk %s.", last(g$rule), pct(last(g$risk), 1))
add("")

## 3. Policy
add("3. WHAT A SCHOOL SHOULD DO")
for (tg in TARGETS) {
  add("   To keep risk under %s:", pct(tg))
  for (s in SCHOOLS) {
    t  <- d3[school == s & target == tg][order(vacc_lab)]
    ok <- t[!is.na(days)]
    mn <- t[label == MIN_POLICY]
    add("     %s: %s; minimal quarantine is enough %s.", s,
        if (nrow(ok)) sprintf("achievable from %s vaccination", as.character(ok$vacc_lab[1]))
        else "not achievable at any tested vaccination level",
        if (nrow(mn)) sprintf("from %s vaccination", as.character(mn$vacc_lab[1]))
        else "at no tested level")
  }
}
add("")

## 4. Optimal low-risk days
add("4. LOW-RISK QUARANTINE")
for (s in SCHOOLS) {
  o <- d4[school == s]
  add("   %s: the best low-risk quarantine is usually about %g days.", s, median(o$opt_low))
  add("     It is not worth it (risk drops less than %s) in %s of high/medium settings.",
      pct(MIN_GAIN), pct(mean(o$opt_low == 0), 1))
  add("     On average, risk goes %s -> %s with the best number of days,",
      pct(mean(o$risk0), 1), pct(mean(o$risk_opt), 1))
  add("     and only to %s even with the longest low-risk quarantine.", pct(mean(o$risk_max), 1))
}
add("")

## 5. Size
add("5. SIZE OF OUTBREAKS")
for (s in SCHOOLS) {
  b <- sims[school == s & q_high == min(HIGHS) & q_med == 0 & q_low == 0,
            .(share = mean(size > 50)), by = vacc][order(vacc)]
  add("   %s, minimal quarantine: chance of more than 50 cases falls from %s at %s to %s at %s vaccination.",
      s, pct(b$share[1], 1), pct(b$vacc[1]), pct(last(b$share), 1), pct(last(b$vacc)))

  op <- opt_pol[school == s][order(vacc)]
  bo <- sims[op, on = .(school, vacc, q_high, q_med, q_low)][
    , .(share = mean(size > 50)), by = vacc][order(vacc)]
  add("     Optimal quarantine (high/medium/low days): %s.",
      paste(sprintf("%s at %s", op$policy, pct(op$vacc)), collapse = "; "))
  add("     With it, the chance of more than 50 cases is %s at %s and %s at %s vaccination.",
      pct(bo$share[1], 1), pct(bo$vacc[1]), pct(last(bo$share), 1), pct(last(bo$vacc)))
}

writeLines(out, file.path(RESULTS_DIR, "conclusions.txt"))
```

```{r print-conclusions, results='asis'}
cat("```text", out, "```", sep = "\n")
```

